Testing Quasar Unification with Clumpy Wind Models

### 2015-11-19 paper.tex
Initial comments on paper.tex

synthetic spectra of simple biconical disc wind models, using our state-of-the-art 
Monte Carlo radiative transfer and photoionization code.
In particular, we investigate whether clumping of the outflow allows us to 
produce synthetic spectra in the rest-frame UV that have the characteristics of quasars. 
We find that a simple treatment of clumping (`microclumping') 
allows for a more realistic X-ray luminosities, while maintaining the ionization 
state necessary for strong BAL features in the rest-frame UV. 

@@

The dense, X-ray heated wind produces strong recombination and collisionally 

@@

Evidence for dense substructures in AGN winds is widespread.
BALQSOs show complex absorption line profiles \citep{ganguly2006, simonhamann2010}
and exhibit variability in these profile shapes \citep{capellupo2011,capellupo2012,capellupo2014}.
AGN generally show variability in X-ray absorption components \citep[e.g.][]{risaliti2002}
and many models for the BLR consist of clumps embedded in an outflow 
\citep{krolik1981, emmering1992, dekool1995, cassidyraine1996}.
Clumping can be caused by magnetic confinement \citep{dekool1995},
or the instabilities inherent to line-driven winds 
\citep{lucysolomon1970,macgregor1979,carlberg1980,owockirybicki1984,owockirybicki1985}.
Additionally, clumping is required to explain the electron scattering wings of emission lines formed
in line-driven hot star winds \citep{hillier1991eswingsmodel}. Complex substructures 
are also produced in simulations of line-driven 
outflows in AGN, although on very different scales to line-driven instabilities 
\citep{PSK2000,PK04,progakurosawa2010,proga2014}.
Nevertheless, clumpy winds offer an observationally motivated and theoretically 
predicted way to lower the ionization state of a plasma, possibly in tandem
with a shielding scenario.

@@

For this study, we use the MCRT code \py\ we have developed to carry out our 
radiative transfer and photoionization simulations in non-local thermodynamic
equilibrium (non-LTE). The code can be model a variety of
disc-wind systems; it has been used with application to accreting white dwarfs 
(Long \& Knigge 2002, hereafter LK02; Noebauer et al. 2010; 
Matthews et al. 2015, hereafter M15), young-stellar objects 
\citep{simmacro2005} and quasars/AGN (H13, H14).\nocite{noebauer, M15, LK02}

The basic premise involves generating photons from an accretion disc
and central object, and computing their radiative transfer through an outflow.
This outflow is discretized into $n_x \times n_z$ cells in a 2.5D
cylindrical geometry with azimuthal symmetry.
First, the code gradually converges on the global temperature and ionization structure.
Once these ``ionization cycles'' are complete, then a series of further iterations
allow one to obtain a detailed synthetic spectrum.
LK02 provides the basic description of the original code; improvements, 
see especially \cite{simmacro2005}, H13 and M15, have been 
described in subsequent reports.  
%A full description of the code will be published elsewhere; 
We focus here on the specific changes made to improve the ionization calculation 
and to allow for clumping in the wind.
%These studies contain extensive detail on the code,  so we only briefly describe the key elements of the global ionization calculation and other important aspects.

@@

\subsection{Line transfer}

Our approach to line transfer is based upon the macro-atom implementation developed by 
\cite{lucy2002, lucy2003}, in which the energy flows through the system are described in 
terms of indivisible energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section~\ref{sec:photon_sources}). 
In our case, for reasons of computational efficiency, we adopt the  hybrid macro-atom scheme 
described by M15.
%(an improvement over that used by H14 for our earlier quasar study), 
In this scheme, the energy packets interact with either two-level `simple ions' 
or full `macro-atoms'. This allows one to treat non-LTE line transfer in radiative equilibrium
without approximation for elements that are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple-ions (see M15). In this study,
only H is treated as a macro-atom, because 
we expect recombination to be important
in determining their level populations and resultant line emission, and because we were especially interested in the contribution to AGN spectra of \LA.  
H13 treated all atoms in a two-level approximation.

\subsection{Ionization Scheme}

@@

In order to calculate the photoionization rate, 
we model the SED in a grid cell using the technique described by H13. In this scheme,
the mean intensity, $J_{\nu}$ in a series of $n$ bands is modeled as either a power law or exponential
in frequency $\nu$, with the fit parameters deduced from band-limited radiation field estimators.
% \begin{equation}
% J_{\nu,i}=K_{pl}\nu^{\alpha_{pl}},
% \end{equation}
% for a band $i$, or an exponential 
% \begin{equation}
% J_{\nu,i}=K_{exp} e^{(-h\nu/k T_{exp})}.
% \end{equation}
% Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
% deduced from the band-limited radiation field estimators.
The ionization rate out of ion $j$ can then be written as 
\begin{equation}
R_{j,j+1}(J)= 
\displaystyle{
n_j \left(C_{j} n_e + 
\sum_{band~i=0}^{n}~
{\int_{\nu_i}^{~\nu_i+1} \frac{4 \pi J_{\nu,i}\sigma_j(\nu)}  {h \nu} d\nu}
\right),}
\end{equation}
where $\sigma_j$ is the photoionization cross-section, $n_e$
is the electron density and $C_{j}$
represents the collisional ionization coefficient.
The recombination rate {\em into} ion $j$ is given by 
\begin{equation}
R_{j+1,j}(T_e) = (\alpha^j_{RR} + \alpha^j_{CR}) n_{j+1} n_e,
\end{equation}
where each $\alpha^j$ here is the recombination rate coefficient into the ground state of ion $j$.
The subscripts denote radiative and 
collisional (three-body) recombination, respectively. 
$T_e$ is the electron temperature.
%% We neglect recombination to and from excited states in the simple-ion calculation. 
For simple-ions, we use a dilute Boltzmann equation to calculate 
the population of level $k$ in ionic stage $j$,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
\end{equation}
Here $z_j$ is the partition function of ionic stage $j$,
$T_R$ is the effective radiation temperature,
$E_k$ is the energy difference between level $k$ and the ground state,
and $g_k$ is the statistical weight of level $k$. 
We stress that this approximation is not required for ions
treated as macro-atoms.

@@

We include all free-free, bound-free and bound-bound heating
and cooling processes in the model. For radiative transfer purposes
we treat electron scattering in the Thomson limit, 
but take full account of Compton heating and cooling when
calculating the thermal balance of the plasma (see H13).
Adiabatic cooling is included and represents the only
departure from strict radiative equilibrium, but is insignificant 
in most of the outflow.
% It is dealt with self consistently by the spontaneous destruction
% of $k$-packets with a probability

@@

We use the same atomic data as described by LK02 and since updated by H13 and M15, 
with the addition of direct (collisional) ionization and recombination
data from \cite{dere2007}. 
Photoionization cross-sections are from \textsc{Topbase} \citep{cunto1993} and \cite{vfky}.
%Dielectronic and 
Radiative recombination rate coefficients are taken from 
the \textsc{Chianti} database version 7.0 \citep{dere1997,landi2012}.
We use ground state recombination rates from \cite{badnell2006} where available,
and otherwise default to calculating recombination rates from the Milne
relation. Free-free Gaunt factors are from \cite{sutherland1998}.

@@

The X-ray source is treated as an isotropic sphere at the ISCO,
which emits r-packets according to a power law in flux with index $\alpha_X$, of the form
\begin{equation}
F_X (\nu) = K_X \nu^{\alpha_X}.
\end{equation}
The normalisation, $K_X$ of this power law is such that it 
produces the specified 2-10~keV luminosity, $L_X$.
In addition to the disc and X-ray source, 
the wind is able to reprocess radiation. However, new 
photon packets are not produced in the wind (as in LK02). 
Instead, this reprocessing is dealt with by enforcing strict
radiative equilibrium ({\em modulo} adiabatic cooling; see section~2.3)
via an indivisible energy packet
constraint (see Lucy 2002, M15).

@@

Our previous modelling efforts assumed a smooth outflow, 
in which the density at a given point was determined only by the 
kinematic parameters and mass loss rate. However, as already discussed,
AGN winds exhibit significant substructure -- the outflow is expected to be
{\em clumpy}, rather than smooth, and probably on a variety of scales. 
Implementing a treatment of clumping is challenging, for
two main reasons. First, there are significant 
computational difficulties associated with adequately
resolving and realistically modelling a series of small scale, high density
regions with a MCRT code. Second, the addition of multiple additional degrees of
freedom in the model results in significantly wider parameter space.
Unfortunately, the physical scale lengths and density contrasts 
associated with these parameters cannot be well-constrained from observations.

To allow for clumping in our outflow we adopt a simple approximation
used successfully in stellar wind modelling, known as 
{\em microclumping} \citep{hamann1998,hamann2008}(MORE REFS). 
The key assumption is that typical clump sizes are much smaller than the 
typical photon mean free path, and thus the clumps are 
both geometrically and optically thin. This approach 
allows one to introduce a `volume filling factor', $f_V$. 
The intra-clump medium is assumed to be a vacuum, so the 
density of the clumps is then multiplied by the ``density enhancement'' 
$D=1/f_V$. Opacities, $\kappa$, and emissivities, $\epsilon$, 
can then be expressed as

@@

Clumping the wind has an important effect on the ionization state and has
been proposed as a solution to the so-called `over-ionization problem' in 
disc winds \citep{hamann2013}. This is the main motivation for incorporating microclumping
into our model. This treatment is first-order; it does not adequately
represent the complex substructures and stratifications in ionization
state we expect in AGN outflows. Nevertheless, clumping is clearly
important in these flows, and this parameterization allows a simple estimate
for the effect clumping might have on the ionization state and emergent 
line emission. 
% It is also encouraging that microclumping has been used 
% successfully in fits to O-star wind spectra \citep{hillier1991eswingsmodel}.

@@

Using this prescription, we conducted a limited parameter
search over a 5-dimensional parameter space involving the 
variables $r_{min}$, $\theta_{min}$, $f_V$, $\alpha$ and $R_v$.
The grid points are shown in table 1.
The aim here was to first fix $M_{BH}$ and $\dot{m}$ to their H13 values,
and increase $L_X$ to $10^{45}$~erg~s$^{-1}$ (a more realistic value for a 
quasar of $10^9M_\odot$ and an eddington fraction of $0.2$; see section~\ref{sec:xray}). 
We then evaluated these models based on the following criteria:
\begin{itemize}
\item Does the model avoid over-ionization and thus produce UV absorption lines 
with $BI > 0$ at $\sim20\%$ of viewing angles?
\item Does significant line emission emerge at low inclinations, with $EW\sim40\AA$ in \civ?
\item Do H recombination lines appear in the spectrum, $EW\sim50\AA$ in \la?
\item Do a small subset of BAL angles produce LoBAL features?
\item Does the model compare favourably to quasar composite spectra?
\end{itemize}
In this section, we present one of the most promising models and discuss
the various successes and failures with respect to the above criteria.
This allows us to gain insight into fundamental geometrical 
and physical constraints and assess the potential for unification. 
The full grid, including output synthetic spectra and plots can be found at
\url{jhmatthews.github.io/quasar-wind-grid/}.

@@

Figure~\ref{fig:wind} shows the physical properties of the wind.
The wind rises slowly from the disc at first, with clumped densities
of $n_H \sim 10^{11}~\rm{cm^{-3}}$ close to the disc plane.
The flow then accelerates over a scale length of $R_V=10^{19}~\rm{cm}$
up to a terminal velocity equal to the escape velocity at the streamline base
($\sim10,000~\rm{km~s^{-1}}$). This gradual acceleration means that
the wind exhibits a stratified ionization structure, with low ionization material
in the base of the wind giving way to highly ionized plasma further out.
This is illustrated in figure~\ref{fig:wind} 
by the panels showing the ion fraction $F=n_j/n_{tot}$ of some important ions.
By clumping the wind, we are able to produce the range of ionization states observed
in quasars and BALQSOs, while adopting a realistic $2-10$ keV X-ray luminosity
of $L_{X}=10^{45}~\rm{ergs~s^{-1}}$. Without clumping, this wind would be over-ionized 
to the extent that opacities in e.g., \civ\ would be entirely negligible (see H13).

@@

The UV spectrum is characterised by strong BAL 
profiles at high inclinations ($> 70^\circ$). 
This highlights the first success of our model: 
clumping means the correct ionization state 
is maintained in the presence of strong X-rays, 
allowing large resonance line opacities. 
At the highest inclinations, the 
cooler, low ionization material at the base of the wind
starts to intersect the line of sight. This produces 
multiple absorption lines in species such as \mg,
\al\ and Fe~\textsc{ii}. The potential links to LoBALQSOs and 
FeLoBALQSOs are discussed in section 2.4.
Clearly, the absorption lines in the 
synthetic spectra at BALQSO-like angles do not match the composites; this
is {\em not} due to a failure of our model, but rather because a geometric mean
BALQSO composite spectrum tends to wash out the broad absorption features due to the wide
range of absorption characteristics. To demonstrate this, we show a comparison 
to a {\sl Hubble Space Telescope} STIS spectrum of the high BALnicity BALQSO PG0946+301 
(Arav et al. 2000) in figure \ref{fig:stis}.

The high ionization BAL profiles are often saturated, and the location in velocity space
of the strongest absorption in the profile varies with inclination.
At the lowest inclination BAL sightlines, the strongest absorption occurs at the red edge,
whereas at higher inclinations (and for the strongest BALs)
the trough has a sharp edge at the terminal velocity.
This offers one potential explanation for the wide range of BALQSO absorption
line shapes (see e.g. Trump et al. 2006; Knigge et al 2008, Filiz Ak et al. 2014).
In addition, the line profile shape is strongly dependent 
on the density, ionization and velocity 
profiles intersected by the line of sight. Thus, small tweaks of the velocity
law and angular distributions of streamlines can dramatically alter
the shape of the line.

Nonblack saturation is observed in the absorption troughs of BALQSOs \citep{arav1999a,arav1999b}.
This can be caused either by partial covering of the continuum
source or by scattered contributions to the BAL troughs, necessarily
from an opacity source not cospatial with the BAL forming region.
The scattered light explanation is supported by spectropolarimetry results
\citep{lamy2000}. Our spectra do not show nonblack saturation.
Instead, we find black, saturated troughs at angles $i > 73^\circ$, and the BALs
are non-saturated at lower inclinations. The reasons for this are readily apparent. 
First, the microclumping assumption does not allow for 
porosity in the wind, meaning that it does not naturally produce
a partial covering absorber. To do this, an alternative approach
such as {\em macroclumping} would be required \citep[e.g.][]{surlan2012,hamann2008}.
Second, our wind does not have a significant scattering 
contribution along sightlines which do not pass through the BAL region,
meaning that any scattered component to the BAL troughs is absorbed by line opacity.
This suggests that either the scattering cross-section of the wind must
be increased (with higher mass loss rates or covering factors), or 
that an additional source of electron opacity is required, potentially
in a polar direction above the disc.

@@

The model produces strong emission lines in \civ, \nv\ and \la,
as well as a weak \mg\ line. The shapes and widths of these lines
match the composites fairly well. However, the line-to-continuum ratios 
at low inclinations in our model are significantly weaker than the quasar 
composites. Increasing the density of the outflow, by altering the mass 
loss rate  or velocity law, can produce more line emission.
However, the red wing of the BAL profiles is generally stronger than 
seen in BALQSO spectra and composites. This illustrates a fundamental 
problem with a geometric unification model such as this:
that the line-to-continuum ratios at hight inclinations are
significantly affected by disc foreshortening and limb darkening.
The angular distribution of the disc radiation is clearly
crucially important in determining the emergent line ratios.

@@

>>>>

synthetic spectra of simple biconical disc wind models, produced with our state-of-the-art 
Monte Carlo radiative transfer and photoionization code PYTHON.  Previously, we have shown that we could produce synthetic spectra resembling those of observed AGN for plausible wind models, but only if the X-ray luminosity was limited to about  10$^{42}$ ergs s$^{-1}$.  Here , we investigate the  degree to which  clumping of the outflow allows us to 
produce synthetic spectra in the rest-frame UV that have the characteristics of quasars with X-ray luminosities as large as 10$^{45}$ ergs s$^{-1}$. 
We conclude, perhaps not surprisingly,  that clumping does maintain the ionization state of the wind necessary for strong BAL features in the rest-frame UV at more realistic X-ray luminosities.

@@

In addition, the dense, X-ray heated wind produces strong recombination and collisionally 

@@

Evidence for dense substructures in AGN winds is widespread.
BALQSOs show complex absorption line profiles \citep{ganguly2006, simonhamann2010}
and exhibit variability in these profile shapes \citep{capellupo2011,capellupo2012,capellupo2014}.
AGN generally show variability in X-ray absorption components \citep[e.g.][]{risaliti2002}
and many models for the BLR consist of clumps embedded in an outflow 
\citep{krolik1981, emmering1992, dekool1995, cassidyraine1996}.
Clumping can be caused by magnetic confinement \citep{dekool1995},
or the instabilities inherent to line-driven winds 
\citep{lucysolomon1970,macgregor1979,carlberg1980,owockirybicki1984,owockirybicki1985}.
Additionally, clumping is required to explain the electron scattering wings of emission lines formed
in line-driven hot star winds \citep{hillier1991eswingsmodel}. Complex substructures 
are also produced in simulations of line-driven 
outflows in AGN, although on very different scales to line-driven instabilities 
\citep{PSK2000,PK04,progakurosawa2010,proga2014}.  {\bf I do not believe the AGN simulations have the resolution to produce clumping on the scale of hot star winds}
Nevertheless, clumpy winds offer an observationally motivated and theoretically 
predicted way to lower the ionization state of a plasma, possibly in tandem
with a shielding scenario.

@@

For this study, we use the MCRT code \py\ we have developed to carry out our 
radiative transfer and photoionization simulations in non-local thermodynamic
equilibrium (non-LTE). The code can be used to model a variety of
disc-wind systems; it has been used with application to accreting white dwarfs 
(Long \& Knigge 2002, hereafter LK02; Noebauer et al. 2010; 
Matthews et al. 2015, hereafter M15), young-stellar objects 
\citep{simmacro2005} and quasars/AGN (H13, H14).\nocite{noebauer, M15, LK02}

The code operates as follows:   This outflow is discretized into $n_x \times n_z$ cells in a 2.5D
cylindrical geometry with azimuthal symmetry. From some initial conditions in each cell, the code first calculates the ionization structure of the wind in a series of iterations.  Each iteration in an ``ionization cycle'' consists of generating photons, actually photon packets, from an accretion disc
and central object, and calculating how these photon bundles scatter through the wind (eventually escaping the outflow or hitting the disk). Then updating the ionization structure based on the properties of the radiation field in each cell, and the process is repeated.   Once the ionization structure has converged, the ionization structure is held fixed, and  synthetic spectra are generated at specific inclination angles in a series of "spectral cycles". LK02 provide a more detailed  description of the original code; various improvements have been made since then and are descbied by 
 \cite{simmacro2005}, H13 and M15.  
%A full description of the code will be published elsewhere; 
We focus here on the specific changes made for this study intended to improve the ionization calculation of H and He
and to allow for clumping in the wind.
%These studies contain extensive detail on the code,  so we only briefly describe the key elements of the global ionization calculation and other important aspects.

@@

\subsection{Line transfer treatment}

Our approach to line transfer is based upon the macro-atom implementation developed by 
\cite{lucy2002, lucy2003}, in which the energy flows through the system are described in 
terms of indivisible energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section~\ref{sec:photon_sources}). 
In our case, for reasons of computational efficiency, we adopt the  hybrid macro-atom scheme 
described by M15.
%(an improvement over that used by H14 for our earlier quasar study), 
In this scheme, the energy packets interact with either two-level `simple ions' 
or full `macro-atoms'. This allows one to treat non-LTE line transfer in radiative equilibrium {\bf what does "in radiative equilibrium" here mean}
without approximation for elements that are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple-ions (see M15). In this study,
only H {\bf and He} are treated as a macro-atom, because 
we expect recombination to be important
in determining their level populations and resultant line emission, and because we were especially interested in the contribution to AGN spectra of \LA.  
H13 treated all atoms in a two-level approximation.  {\bf I thought we were also doing He}

\subsection{Ionization treatment (anything but scheme)}

@@

In order to calculate the photoionization rate, 
we model the SED in a grid cell using the technique described by H13. In this scheme,
the mean intensity, $J_{\nu}$ in a series of $n$ bands is modeled as either a power law or exponential
in frequency $\nu$, with the fit parameters deduced from band-limited radiation field estimators.
% \begin{equation}
% J_{\nu,i}=K_{pl}\nu^{\alpha_{pl}},
% \end{equation}
% for a band $i$, or an exponential 
% \begin{equation}
% J_{\nu,i}=K_{exp} e^{(-h\nu/k T_{exp})}.
% \end{equation}
% Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
% deduced from the band-limited radiation field estimators.
The ionization rate out of ion $j$ can then be written as 
\begin{equation}
R_{j,j+1}(J)= 
\displaystyle{
n_j \left(C_{j} n_e + 
\sum_{band~i=0}^{n}~
{\int_{\nu_i}^{~\nu_i+1} \frac{4 \pi J_{\nu,i}\sigma_j(\nu)}  {h \nu} d\nu}
\right),}
\end{equation}
where $\sigma_j$ is the photoionization cross-section, $n_e$
is the electron density and $C_{j}$
represents the collisional ionization coefficient.
The recombination rate {\em into} ion $j$ is given by 
\begin{equation}
R_{j+1,j}(T_e) = (\alpha^j_{RR} + \alpha^j_{CR}) n_{j+1} n_e,
\end{equation}
where each $\alpha^j$ here is the recombination rate coefficient into the ground state of ion $j$.
The subscripts denote radiative and 
collisional (three-body) recombination, respectively. 
$T_e$ is the electron temperature.
%% We neglect recombination to and from excited states in the simple-ion calculation. 
For simple-ions, we use a dilute Boltzmann approximation to calculate 
the population of level $k$ in ionic stage $j$,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
\end{equation}
Here $z_j$ is the partition function of ionic stage $j$,
$T_R$ is the effective radiation temperature,
$E_k$ is the energy difference between level $k$ and the ground state,
and $g_k$ is the statistical weight of level $k$. 
We stress that this approximation is not required for ions
treated as macro-atoms. {\bf It seems a bit odd that you explain what we do with simple ions our old approach, but don't say what is done for macro-atoms,.  You only say it is not the other.}

@@

We include  free-free, bound-free and bound-bound heating
and cooling processes in the model. For radiative transfer purposes
we treat electron scattering in the Thomson limit, 
but take full account of Compton heating and cooling when
calculating the thermal balance of the plasma (see H13).
Adiabatic cooling is included, but is insignificant 
in most of the outflow.
% It is dealt with self consistently by the spontaneous destruction
% of $k$-packets with a probability

@@

We use the same atomic data  described by LK02 as updated by H13 and M15, 
with the addition of direct (collisional) ionization and recombination
data from \cite{dere2007}. 
Photoionization cross-sections are from \textsc{Topbase} \citep{cunto1993} and \cite{vfky}.
%Dielectronic and 
Radiative recombination rate coefficients are taken from 
the \textsc{Chianti} database version 7.0 \citep{dere1997,landi2012}.
We use ground state recombination rates from \cite{badnell2006} where available,
and otherwise default to calculating recombination rates from the Milne
relation. Free-free Gaunt factors are from \cite{sutherland1998}.

@@

The X-ray source is treated as an isotropic sphere at the ISCO,
which emits r-packets according to a power law in flux with index $\alpha_X$, of the form
\begin{equation}
F_X (\nu) = K_X \nu^{\alpha_X}.
\end{equation}
The normalisation, $K_X$ of this power law is such that it 
produces the specified 2-10~keV luminosity, $L_X$.
In addition to the disc and X-ray source, 
the wind is able to reprocess radiation. However, new 
photon packets are not produced in the wind (as in LK02). 
Instead, this reprocessing is dealt with by enforcing strict
radiative equilibrium ({\em modulo} adiabatic cooling; see section~2.3)
via an indivisible energy packet
constraint (see Lucy 2002, M15).  {\bf I confess, I do not understand what is being said here.  I would like to understand it better, and I would like it explained better.  How do we avoid, generating for example recombination photons from non-macro atoms.}

@@

Our previous modelling efforts, we have assumed a smooth outflow, 
in which the density at a given point was determined only by the 
kinematic parameters and mass loss rate. However, as already discussed,
AGN winds exhibit significant substructure -- the outflow is expected to be
{\em clumpy}, rather than smooth, and probably on a variety of scales. 
Implementing a treatment of clumping is challenging, for
two main reasons. First, there are significant 
computational difficulties associated with adequately
resolving and realistically modelling a series of small scale, high density
regions with a MCRT code. Second, the addition of multiple additional degrees of
freedom in the model results in significantly wider parameter space.
Unfortunately, the physical scale lengths and density contrasts 
associated with these parameters are not well-constrained from observations.  {\bf Reverse the order of these difficulties.  The most important is that it is not clear what kind of clumping an AGN wind has; the computational problem is a secondary one.}

Given the lack of knowledge about the actual type of clumping, we have adopt a simple approximation
used successfully in stellar wind modelling, known as 
{\em microclumping} \citep{hamann1998,hamann2008}(MORE REFS), which minimizes the computational difficulties. 
The underlying assumption in microclumping is that clump sizes are much smaller than the 
typical photon mean free path, and thus the clumps are 
both geometrically and optically thin. This approach 
allows one to introduce a `volume filling factor', $f_V$. 
The intra-clump medium is assumed to be a vacuum, so the 
density of the clumps is then multiplied by the ``density enhancement'' 
$D=1/f_V$. Opacities, $\kappa$, and emissivities, $\epsilon$, 
can then be expressed as

@@

Clumping the wind has an important effect on the ionization state and has
been proposed as a solution to the so-called `over-ionization problem' in 
disc winds \citep{hamann2013}. This is the main motivation for incorporating microclumping
into our model. This treatment is first-order; it does not adequately
represent the complex substructures and stratifications in ionization
state we expect in AGN outflows. Nevertheless,  this parameterization allows simple estimates
of the effect clumping might have on the ionization state and emergent 
line emission. 
% It is also encouraging that microclumping has been used 
% successfully in fits to O-star wind spectra \citep{hillier1991eswingsmodel}.

@@

Using this prescription, we conducted a limited parameter
search over a 5-dimensional parameter space involving the 
variables $r_{min}$, $\theta_{min}$, $f_V$, $\alpha$ and $R_v$.
The grid points are shown in table 1.
The aim here was to first fix $M_{BH}$ and $\dot{m}$ to their H13 values,
and increase $L_X$ to $10^{45}$~erg~s$^{-1}$ (a more realistic value for a 
quasar of $10^9M_\odot$ and an eddington fraction of $0.2$; see section~\ref{sec:xray}). 
We then evaluated these models based on the following criteria:
\begin{itemize}
\item Does the model avoid over-ionization and thus produce UV absorption lines 
with $BI > 0$ at $\sim20\%$ of viewing angles?
\item Do the synthetic spectra show  line emission emerge at low inclinations, with $EW\sim40\AA$ in \civ?
\item Do H recombination lines appear in the spectra, $EW\sim50\AA$ in \la?
\item Do the spectra  produce LoBAL features at a  small subset of BAL angles?
\item Does the model spectra resemble quasar composite spectra?
\end{itemize}
In this section, we present one of the most promising models and discuss
the various successes and failures with respect to the above criteria.
This allows us to gain insight into fundamental geometrical 
and physical constraints and assess the potential for unification. 
The full grid, including output synthetic spectra and plots can be found at
\url{jhmatthews.github.io/quasar-wind-grid/}.

{\bf I still think we need to avoid using model to be a generic term to describe the outflow and the spectra.  If you mean the spectra use "model spectra" or "synthetic spectra'.  Note - the whole question of the model grid and the selection of one specific instantiation is not dealt with well.}

@@

Figure~\ref{fig:wind} shows the physical properties of the wind.
The wind rises slowly from the disc at first, with clumped densities  {\bf clumped densities is jargon, use perhaps density of the clumps in the wind}
of $n_H \sim 10^{11}~\rm{cm^{-3}}$ close to the disc plane.
The flow then accelerates over a scale length of $R_V=10^{19}~\rm{cm}$
up to a terminal velocity equal to the escape velocity at the streamline base
($\sim10,000~\rm{km~s^{-1}}$). This gradual acceleration results in
a wind that exhibits a stratified ionization structure, with low ionization material
in the base of the wind giving way to highly ionized plasma further out.
This is illustrated in figure~\ref{fig:wind} 
by the panels showing the ion fraction $F=n_j/n_{tot}$ of some important ions.
With a clumpted wind, we are able to produce the range of ionization states observed
in quasars and BALQSOs, while adopting a realistic $2-10$ keV X-ray luminosity
of $L_{X}=10^{45}~\rm{ergs~s^{-1}}$. Without clumping, this wind would be over-ionized 
to the extent that opacities in e.g., \civ\ would be entirely negligible (see H13).

@@

The UV spectrum is characterised by strong BAL 
profiles at high inclinations ($> 70^\circ$). 
This highlights the first success of our model: 
clumping allows the correct ionization state 
to be maintained in the presence of strong X-rays, 
resulting in large resonance line opacities. 
At the highest inclinations, the 
cooler, low ionization material at the base of the wind
starts to intersect the line of sight. This produces 
multiple absorption lines in species such as \mg,
\al\ and Fe~\textsc{ii}. The potential links to LoBALQSOs and 
FeLoBALQSOs are discussed in section 2.4.
Clearly, the absorption lines in the 
synthetic spectra at BALQSO-like angles do not match the composites; this
is {\em not} due to a failure of our model, but rather because a geometric mean
BALQSO composite spectrum tends to wash out the broad absorption features due to the wide
range of absorption characteristics. To demonstrate this, we show a comparison 
to a {\sl Hubble Space Telescope} STIS spectrum of the high BALnicity BALQSO PG0946+301 
(Arav et al. 2000) in figure \ref{fig:stis}.

The high ionization BAL profiles are often saturated, and the location in velocity space
of the strongest absorption in the profile varies with inclination.
At the lowest inclination BAL sight lines, the strongest absorption occurs at the red edge,
whereas at higher inclinations (and for the strongest BALs)
the trough has a sharp edge at the terminal velocity.
This offers one potential explanation for the wide range of BALQSO absorption
line shapes (see e.g. Trump et al. 2006; Knigge et al 2008, Filiz Ak et al. 2014).
In addition, the line profile shape is strongly dependent 
on the density, ionization and velocity 
profiles intersected by the line of sight. Thus, small tweaks of the velocity
law and angular distributions of streamlines can dramatically alter
the shape of the line.

Non-black saturation is observed in the absorption troughs of BALQSOs \citep{arav1999a,arav1999b}.
This is usually explained either as  partial covering of the continuum
source or by scattered contributions to the BAL troughs, necessarily
from an opacity source not co-spatial with the BAL forming region.
The scattered light explanation is supported by spectropolarimetry results
\citep{lamy2000}. Our spectra do not show nonblack saturation.
Instead, we find black, saturated troughs at angles $i > 73^\circ$, and the BALs
are non-saturated at lower inclinations. The reasons for this are inherent in the construction of our model. 
First, the microclumping assumption does not allow for 
porosity in the wind, meaning that it does not naturally produce
a partial covering absorber. To allow this, an alternative approach
such as {\em macroclumping} would be required \citep[e.g.][]{surlan2012,hamann2008}.
Second, our wind does not have a significant scattering 
contribution along sightlines which do not pass through the BAL region,
meaning that any scattered component to the BAL troughs is absorbed by line opacity.
This suggests that either the scattering cross-section of the wind must
be increased (with higher mass loss rates or covering factors), or 
that an additional source of electron opacity is required, potentially
in a polar direction above the disc.  {\bf it sounds like what you are suggesting is a low density inner flow, our domain 2 stuff actually permits ... or indeed one might be able to produce this within SV is you adopted a different mass loss with radius assumption}

@@

The model produces strong emission lines in \civ, \nv\ and \la,
as well as a weak \mg\ line. The shapes and widths of these lines
match the composites fairly well. However, the line-to-continuum ratios 
at low inclinations in our model are significantly weaker than the quasar 
composites. Increasing the density of the outflow, by altering the mass 
loss rate  or velocity law, can produce more line emission.
However, the red wing of the BAL profiles is generally stronger than 
seen in BALQSO spectra and composites. This illustrates a fundamental 
problem with a geometric unification model such as this:
that the line-to-continuum ratios at high inclinations are
significantly affected by disc foreshortening and limb darkening.
The angular distribution of the disc radiation is clearly
crucially important in determining the emergent line ratios.  {\bf This is a bit oddly explained. Our problem is at low inclination.  Note that I am a bit concerned with Fig 3; the model spectra and the observations look very different.}

@@

{\bf I have to say I find this hard to understand.  Are you trying to say that in CVs which we are know are thin disks with winds, the EW of the red-wing of emission lines increases with inclination angle ... which I think may be true, and so one would expect that in a unification model for AGN  it should do the same?  And the fact that it does not, suggests that something else is going on in AGN.  I also still think that you should move this discussion later in the paper and not bury it here.}
